---
title: "FREQUENCY RANGE DETECTION OF Eudynamys scolopaceus"
author: "OJO SEGUN ADEWALE"
date: "APRIL 4, 2019"
output: 
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
library("warbleR")
library("tuneR")
library("Rraven")
library("devtools")
library("knitr")
library("magick")
setwd("C:/Users/PC/Desktop/MB2_PRESENTATION")
getwd()
```

## An Overview

 - Frequency Range Detection
 
   + Acoustic Monitoring of (Eudynamys scolopaceus) sound
   
   + Getting the frequency range of acoustic signals from Spectrum
   
   + Filtering of the needed specie's sound from Xeno-canto using three distinct feature
   
   + Detection of downloaded sound in directory automatically and conversion to .wav files for easy analysis
   
   + Followed by Feature extraction and analysis.


## Introduction
 - Asian Koel  (Eudynamys scolopaceus)
  + Mainly resident breeder in Tropical Southern Asia
    
  + They first arrived in Singapore in the 1980s 
    
  + A crafty avian parasite, shy bird, more heard than seen
    
  + They don't build nest but lays Egg in the nest of other birds (crow's nest)
   
  + They push eggs of the host to accommodate their own
   
  + 39- 46cm in Length, Eyes are bright red
   
  + Lives in light woodland , coastal mangrove's edges
   
  + Feeds on caterpillars and insect for protein
    
 


## Load needed Packages

 + library("warbleR")
 + library("tuneR")
 + library("Rraven")
 + library("devtools")
 + library("knitr")
 + library("magick")

## Querying and checking out the Structure of all Recordings of Eudynamys species present.
```{r,echo=TRUE}
##Query  xeno-canto for all recordings of Eudynamys
Asian<- querxc(qword="Eudynamys", download= FALSE)
```


##Check out the structure of resulting the data frame
```{r,echo=TRUE}
str(Asian)
```

##Query xeno-canto for all Eudynamys scolopaceus recordings
```{r,echo=TRUE}
Asian.koel<-querxc(qword="Eudynamys scolopaceus", download = FALSE)
```

##Check out the structure of resulting the data frame
```{r,echo=TRUE}
str(Asian.koel)
```

## Geographical distribution of the Asian Koel
```{r,echo=TRUE}
#search and query and get the numbers and how the species are spread accross
X <- quer_xc("Eudynamys scolopaceus", download = FALSE) ##Eudynamys-scolopaceus
```


```{r,echo=TRUE}
## Output of view X, displayed in the next window automatically
View(X) # Including available sounds,location, Recordist e.t.c
## write X into the Directory as csv
write.csv(X, "C:/Users/PC/Desktop/MB2_PRESENTATION/X.csv")
xcmaps(X)
xcmaps(X, img = FALSE, it = "tiff") 
# The generated Map can be found in The working Directory
```

## Read image xcmaps from Directory with magick function
```{r,echo=TRUE}
xcmaps<- image_read("C:/Users/PC/Desktop/MB2_PRESENTATION/Map of Eudynamys scolopaceus recordings.jpeg")
print(xcmaps) ## Shows Geographical distribution of specie
```

## Existing levels of Signal for the Specie
```{r,echo=TRUE}
##Query to get the number of signal that exist in xeno-canto for Asian-koel
levels(Asian.koel$Vocalization_type)
```

##Query to get the number of recordings per signal type
```{r,echo=TRUE}
##The output cannot be seen verywell because it's on the right hand side 
##but can be seen in the console 
table(Asian.koel$Vocalization_type)
```

## Query meta data to select the best signal using three filter; First by quality
```{r,echo=TRUE}
# Quality 
Asian.koel<-Asian.koel[Asian.koel$Quality=="A",]
nrow(Asian.koel)
```


## Filter secondly by Signal and thirdly by Location 
```{r,echo=TRUE}
##Second by signal type
Asian.koel.call<-Asian.koel[grep("call", Asian.koel$Vocalization_type, ignore.case = TRUE), ]
nrow(Asian.koel.call)

###Third by locality
Asian.koel.LS<-Asian.koel.call[grep("Dinigat Island",Asian.koel.call$Locality, ignore.case = FALSE),]
```

## Check the Structure of resulting data frame
```{r,echo=TRUE}
str(Asian.koel.LS)
```

## Download the recording
```{r,echo=TRUE}
querxc(X=Asian.koel.LS)
```


## Read file and convert to .wav file
```{r,echo=TRUE}
# Query and download  Xeno-Canto for metadata using genus and species as keywords 
##asian.koel <- quer_xc(qword = "nr:446387", download = TRUE, pb = FALSE)
r <- readMP3("Eudynamys-scolopaceus-446387.mp3")  ## MP3 file in working directory
writeWave(r,"tmp.wav",extensible=FALSE)

# Simultaneously lower sampling rate to speed up down stream analyses
mp32wav(samp.rate = 22.05, pb = FALSE)


ad <- auto_detec(wl = 200, threshold = 3.5, ssmooth = 1200, bp = c(4, 9.6), 
                 pb = FALSE,  mindur = 0.1, maxdur = 0.25, img = FALSE)
```


```{r,echo=TRUE}
snr <- sig2noise(ad, pb = FALSE, mar = 0.05)

ad <- snr[rank(-snr$SNR) < 60, ]
# A look at the subset of signals to be analyzed
catalog(ad, nrow = 5, ncol = 4, mar = 0.01, flim = c(3.5, 10.5), 
        labels = "selec", pal = reverse.heat.colors, pb = FALSE)
open_wd() ## check wd, produces two catalogs
```

##Read image catalog1 from Directory with magick function
```{r,eval = TRUE, echo=FALSE}
catalog1<- image_read("C:/Users/PC/Desktop/MB2_PRESENTATION/Catalog_p1.jpeg")
print(catalog1)
```

##Read image catalog2 from Directory with magick function
```{r,eval = TRUE, echo=FALSE}
catalog2<- image_read("C:/Users/PC/Desktop/MB2_PRESENTATION/Catalog_p2.jpeg")
print(catalog2)
```

## Freq Range detect.for selection 2 only
```{r,echo=TRUE}
w2 <- readWave(as.character(ad$sound.files[2]), from = ad$start[2], 
               to = ad$end[2], units = "seconds")
freq_range_detec(w2, bp = c(2, 9.6), fsmooth = 1, ovlp = 95, 
                 wl = 200, threshold = 8)## Pre-selection
```

## Freq Range detection for all selection
```{r,echo=TRUE}
for(i in 1:nrow(ad))
{
  wv <- readWave(as.character(ad$sound.files[i]), from = ad$start[i], 
                 to = ad$end[i], units = "seconds")
  freq_range_detec(wv, bp = c(2, 12), fsmooth = 1, ovlp = 95, wl = 200, 
                   threshold = 8, flim = c(0, 13), 
                   main = paste("selection", i))
  Sys.sleep(0.8)
} # All 39 .wavfiles check (WD)- this folder (Eudynamys-scolopaceus_ojo_files)
```

## Calculate for all selection using freq_range
```{r,echo=TRUE}
## Output displayed in the next window automatically.
fr_ad <- freq_range(X = ad, bp = c(2, 12), fsmooth = 0.001, ovlp = 95, 
                    wl = 200, threshold = 10, img = FALSE)
View(fr_ad) # Contains start, end, SNR,Top&Bottom Freq.
## write (fr_ad) into the Directory as csv
write.csv(fr_ad, "C:/Users/PC/Desktop/MB2_PRESENTATION/fr_ad.csv")
```

## Dropping of NAs based on the Threshold
```{r,echo=TRUE}
# Check rows that have NAs
row.has.na <- apply(fr_ad, 1, function(x){any(is.na(x))})

# summation of number of rows with NA values
sum(row.has.na)

# Dropping of rows with NA values
fr_ad.filtered <- fr_ad[!row.has.na,]
```

## Measuring of over 28 Acoustic Parameters in the sound files
```{r,echo=TRUE}
#using the range as Band pass filter
sp <- specan(X = fr_ad.filtered, wl = 200, bp = "frange", ovlp = 90, pb = FALSE, 
             threshold = 15)
View(sp)## Output displayed in the next window automatically
##write (sp) into the Directory as csv
write.csv(sp, "C:/Users/PC/Desktop/MB2_PRESENTATION/sp.csv")
```
## Conclusion
 + According to the IBA, Asian koel is one of the LCB in terms threat
 + Dinagat Island is one of the 117key conservation site in Philippines
 + The SNR divides the Mean Amplitude of the Signal by the Mean Amplitude of background noise adjacent to the signal
 + Detection of these variouse Signals across the subset(catalog 1&2) is quite helpful and parameter adjustment can     as well improve the detection
 + Freq_range helps to get the needed range for each of the 39 files however rows with NAs were dropped
 + Using the Freq_range as a band filter generates over 28 acoustic parameters for the remaining 35 .wav files


## For further analysis and information about the Specie and package used   
https://www.researchgate.net/profile/Jose_Don_De_Alban/publication
/281290212_Mts_Kambinliw_Redondo_Loreto_Dinagat_Island_Surigao_del
_Norte_Integrating_Forest_Conservation_with_Local_Governance_Technical
_Report/links/55e035d208aecb1a7cc22a83.pdf

https://cran.r-project.org/web/packages/warbleR/vignettes/warbleR_workflow
_01.html#next_vignette:_visual_inspection_and_signal_classification

https://cran.r-project.org/web/packages/magic/index.html

Link to Github;
https://github.com/OJCITY/Rproject

Thank you all for  your time